Simulating High-Dimensional Multivariate Data

### 2021-02-24 010-background.Rmd

```{r ch010-background, include=FALSE, cache=FALSE}
devtools::load_all()
library(tidyverse)
```

@@

We illustrate our methodology using a well-studied Breast Invasive Carcinoma (BRCA) data set housed in The Cancer Genome Atlas (TCGA; see Acknowledgments).
For simplicity, we only consider high expressing genes. In turn, we begin by filtering to retain the top `r d`% of highest-expressing genes (in terms of median expression) of the 20,502 gene measurements from $N=`r nrow(brca)`$ patients' tumor samples. This gives a great number of pairwise dependencies among the marginals (specifically, $`r choose(d,2)`$ correlation parameters). Table \@ref(tab:ch010-realDataTab) displays RNA-seq counts for three selected high-expressing genes for the first five patients' breast tumor samples.

@@

small_brca <- brca %>%
  select(all_of(gene_sample)) %>%
  head(n = num_patients) %>%
  as_tibble(rownames = "Patient ID") %>%
  mutate(`Patient ID` = str_sub(`Patient ID`, end = 12))

>>>>

```{r ch010-background, include=FALSE, cache=FALSE}
devtools::load_all()
library(tidyverse)
d <- 1000
```

@@

We illustrate our methodology using a well-studied Breast Invasive Carcinoma (BRCA) data set housed in The Cancer Genome Atlas (TCGA; see Acknowledgments).
For simplicity, we only consider high expressing genes. In turn, we begin by filtering to retain the top `r d`% of highest-expressing genes (in terms of median expression) of the 20,502 gene measurements from $N=`r nrow(example_brca)`$ patients' tumor samples. This gives a great number of pairwise dependencies among the marginals (specifically, $`r choose(d,2)`$ correlation parameters). Table \@ref(tab:ch010-realDataTab) displays RNA-seq counts for three selected high-expressing genes for the first five patients' breast tumor samples.

@@

small_brca <- example_brca %>%
  select(all_of(gene_sample)) %>%
  head(n = num_patients) %>%
  as_tibble(rownames = "Patient ID") %>%
  mutate(`Patient ID` = str_sub(`Patient ID`, end = 12))

### 2021-02-24 020-simulation-algorithm.Rmd

```{r ch020-preview, echo = FALSE, eval = FALSE}
bookdown::preview_chapter('020-simulation-algorithm.Rmd')
```

@@

## `bigsimr::rvec` for random vector generation  {#rand-vec-gen}

- Discuss how the HD NORTA defective problem and the scalability issues lead to our algorithm
We now describe `bigsimr::rvec`, our algorithm to generate random vectors, which mirrors the classical NORTA algorithm above with some modifications for rank-based dependency matching:

@@

The "mapping step" takes advantage of the closed-form relationships between $\rho_S$ and $\tau$ with $\rho_P$ for bivariate normal random variables via Equations \@ref(eq:convertKendall) or \@ref(eq:convertSpearman), respectively (implemented as `bigsimr::cor_covert`).

@@

$$
\begin{aligned}
\Omega_P &= \{ R_P : R_P \,  is \, a \,  correlation \,  matrix \,  for \,  X \} \\
\Omega_K &= \{ R_K : R_K \,  is \,  a \,  Kendall \,  \tau  \, matrix \,  for \,  X \} \\
\Omega_S &= \{ R_S : R_S \,  is \,  a \,  Spearman \,  correlation \,  matrix \,  for \,  X\} \\
\end{aligned}
$$

@@

The final step implements a high-performance NORTA algorithm.
Specifically, step 3i uses either the efficient multi-core multivariate normal simulator `mvnfast`[@Fasiolo2016] or Google's JAX python library for GPU acceleration of the Cholesky factorization and matrix multiplication and steps 3ii, 3iii are parallelized over CPU cores.

>>>>

```{r ch020-preview, eval=FALSE}
bookdown::preview_chapter('020-simulation-algorithm.Rmd')
```

@@

## `rvec` for random vector generation {#rand-vec-gen}

- Discuss how the HD NORTA defective problem and the scalability issues lead to our algorithm
We now describe `rvec`, our algorithm to generate random vectors, which mirrors the classical NORTA algorithm above with some modifications for rank-based dependency matching:

@@

The "mapping step" takes advantage of the closed-form relationships between $\rho_S$ and $\tau$ with $\rho_P$ for bivariate normal random variables via Equations \@ref(eq:convertKendall) or \@ref(eq:convertSpearman), respectively (implemented as `cor_covert`).

@@

$$
\begin{aligned}
\Omega_P &= \{ R_P : R_P \textrm{ is a correlation matrix for } X \} \\
\Omega_K &= \{ R_K : R_K \textrm{ is a correlation matrix for } X \} \\
\Omega_S &= \{ R_S : R_S \textrm{ is a correlation matrix for } X \} \\
\end{aligned}
$$

@@

The final step implements a high-performance NORTA algorithm. Specifically, step 3i uses a fast multi-core multivariate normal simulator, and steps 3ii, 3iii are parallelized over CPU cores.

### 2021-02-24 050-examples.Rmd

```{r ch050-LoadLib, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide", cache=F}
## knitr::opts_chunk$set(echo = FALSE, cache = FALSE)
# reticulate::use_condaenv("bigsimr")
library(bigsimr)
library(tidyverse)
cores <- as.integer( parallel::detectCores() - 1 )
set.seed(10142020)
```

```{r ch050-preview, echo = FALSE, eval = FALSE}
bookdown::preview_chapter('050-examples.Rmd')
```

@@

```{r ch050-readBRCA, echo=FALSE, eval=TRUE, cache=F}
d <- 1000
brcaRDS <- paste0( "data/brca", d, ".rds" )
brca <- readRDS(file = brcaRDS)
## smaller set for prototyping
## d=5
## brca <- brca[ , 1:d]
```

@@

With this goal in mind, we first estimate the desired correlation matrix using the fast implementation provided by `bigsimr`.:

```{r ch050-estRhoBRCA, echo=TRUE, eval=TRUE, cache=F}
## Estimate Spearman's correlation on the count data
type <- 'spearman'
R_S <- bigsimr::cor_fast(brca, method = type)
```

Next, we estimate the marginal parameters. 
We use the method of moments (MoM) to estimate the marginal parameters for the multivariate negative binomial model.
The marginal distributions are from the same probability family (NB) yet are heterogeneous in terms of the parameters probability and size $(p_i, n_i)$ for $i,\ldots,d$.
The functions below support this estimation for later use in `bigsimr::rvec`.

<!-- 
TYPESTE MoM estimators here.
your comment -->

```{r ch050-nbHelpers, echo=TRUE}
make_nbinom_alist <- function(sizes, probs) {
  lapply(1:length(sizes), function(i) {
    substitute(qnbinom(size = s, prob = p), 
               list(s = sizes[i], p = probs[i]))
  })
}
## make_nbinom_alist(c(20, 21, 22), c(0.3, 0.4, 0.5))
nbinom_mom <- function(x) {
  m <- mean(x)
  s <- sd(x)
  s2 <- s^2
  p <- m/s2
  r <- m^2 / (s2 - m)
  c(r, p)
}
```

@@

```{r ch050-estMargins, echo = TRUE, eval = TRUE, cache=F}
sizes <- apply( unname(as.matrix(brca)), 2, nbinom_mom )[1, ]
probs <- apply( unname(as.matrix(brca)), 2, nbinom_mom )[2, ]
## construct margins
nb_margins <- make_nbinom_alist(sizes, probs)
```

@@

```{r ch050-cor-check, echo=TRUE, eval=TRUE, cache = TRUE}
## 1. Mapping step first
R_X <- cor_convert(rho = R_S, from = type, to = 'pearson')
## 2a. Check admissibility
is_valid_correlation <- function(rho) {
  all(diag(rho) == 1) && matrixcalc::is.positive.semi.definite(rho)
}
is_valid_correlation( R_X )

## 2b. compute nearest correlation
if ( ! is_valid_correlation( R_X ) ) {
    ## system.time( R_Xtilde <- cor_nearPSD( G = R_X )  )
    R_Xtilde <- cor_nearPSD( G = R_X )
    ## Quantify the error
    targets <- R_X[ lower.tri( R_X ) ]
    approximates <- R_Xtilde[ lower.tri( R_Xtilde ) ]
    ## R_Xtilde[1:5, 1:5]
    ## norm( x = R_X - R_Xtilde, type = "F" )
    R_X  <- R_Xtilde
}
## qplot( x = targets, y = approximates ) + geom_abline(slope=1,intercept = 0)
## qplot( x = targets - approximates ) + geom_abline(slope=1,intercept = 0)
summary(targets - approximates)
```

While the exact dxd Spearman correlation matrix is not strictly admissible in our scheme (seen by the non-PSD result above), the approximation is of sufficient quality for most applications with a maximum pairwise of `r max( targets - approximates)` and average error of `r mean( targets - approximates)`.

```{r ch050-runBRCA-echo, echo=TRUE, eval=FALSE, cache = FALSE}
sim_nbinom <- rvec(10000, R_X, nb_margins, ensure_PSD = FALSE, cores = cores) 
colnames(sim_nbinom) <- names(brca)
```

```{r ch050-runBRCA, echo=FALSE, results=FALSE, cache = FALSE}
brcaSimRDS <- paste0( "results/brca", d, "sim.rds" )
if( !file.exists(brcaSimRDS) ) {
    ## Set the number of random vectors
    n <- 10000
    ## construct margins
    ## run sims
    set.seed(44)
    sim_nbinom <- rvec(10000, R_X, nb_margins, ensure_PSD = FALSE, cores = cores)
    saveRDS( sim_nbinom, brcaSimRDS )
}
sim_nbinom <- readRDS( brcaSimRDS )
```

@@

```{r ch050-makeSimDataFig, echo = FALSE, eval=T, cache=F, message=FALSE, warning=FALSE}
if( !file.exists("fig/ch050-simDataFig.png") ) {
    exampleGenes <- readRDS(file = './data/exampleGenes.rds')
    p <- GGally::ggpairs(
                     data = as.data.frame( sim_nbinom[ ,exampleGenes] ), 
                     upper = list(
                         continuous = GGally::wrap('cor', method = "spearman")
                     )) + theme_bw()
    ## ggsave("fig/ch050-simDataFig.png", plot = p, width = 5, height = 4, dpi = 300)
    ggsave("fig/ch050-simDataFig.png", plot = p, dpi = 300) 
}
```

```{r ch050-simDataFig, echo=F, fig.align='center', fig.cap='Simulated data for three selected high-expressing genes generally replicates the estimated data structure. The data do not exhibit outlying points, but do possess the desired Spearman correlations, central tendencies, and discrete values.'}
knitr::include_graphics('fig/ch050-simDataFig.png')
```

@@

```{r ch050-makeFigBRCA, echo=F, eval=T, cache=F}
if( !file.exists("fig/ch050-figBRCA.png") ) {
    ## estimate rho from sims and plot
    R_S_hat <- cor_fast(sim_nbinom, method = type)
    ## triplot
    ## png("fig/ch050-figBRCA.png", width = 4, res = 100, units = 'in')
    png("fig/ch050-figBRCA.png", res = 300, width = 5, height = 4, units = 'in')
    par(mfrow=c(1,3))
    plot(R_S[lower.tri(R_S)], R_S_hat[lower.tri(R_S_hat)],
         xlim = c(-1, 1), ylim = c(-1, 1),
         pch = 1, cex = 1, col = rgb(0, 0, 0, 0.05),
         xlab = "Target Correlation", ylab = "Estimated Correlation")
         ## main = "Spearman Cor. Matching")
    abline(a = 0, b = 1, col="red", lty="dashed")
    ## estimate prob,size from sims and plot
    nbinom_size_prob_hat <- apply(sim_nbinom, 2, nbinom_mom)
    plot(sizes, nbinom_size_prob_hat[1,],
         pch = 1, cex = 1, col = rgb(0, 0, 0, 1),
         xlab = "Target Size", ylab = "Estimated Size")
         ## main = "Marginal NB Size Matching")
    abline(a=0, b=1, col="red", lty="dashed")
    plot(log(probs), log(nbinom_size_prob_hat[2,]),
         pch = 1, cex = 1, col = rgb(0, 0, 0, 1),
         xlab = "Target Log Probability", ylab = "Estimated Log Probability")
         ## main = "Marginal NB Prob Matching")
    abline(a=0, b=1, col="red", lty="dashed")
    dev.off()
    par(mfrow=c(1,1))
}
```

```{r ch050-figBRCA, echo=F, out.width = '80%', fig.align='center', fig.cap="Simulated random vectors from a multivariate negative binomial replicate the estimated structure from an RNA-seq data set. The dashed red lines indicated equality between estimated parameters from simulated data (vertical axes) and the specified target parameters (horizontal axes)."}
knitr::include_graphics('fig/ch050-figBRCA.png')
```

@@

```{r ch050-densityEvaluation, include=FALSE, eval=TRUE, cache=F}
d <- ncol(sim_nbinom)
B <- nrow(sim_nbinom)
threshold <- rep(1, d)
pHat <- mean(apply(sim_nbinom, 1, function(X, y0 = threshold) {
  all(X >= y0)
}))
```

where ${\bf Y^{(b)} }$ is the $b^{th}$ simulated vector from a total of $B$ simulation replicates and $I ( )$ is the indicator function.
For example, we can estimate from our $B=10,000$ simulated vectors that the probability of all genes expressing (i.e., ${\bf y}_i \geq 1, \forall \; i$) is $`r pHat`$.

```{r ch050-densityEvaluationECHO, echo=TRUE, eval=TRUE, cache=F}
d <- ncol(sim_nbinom)
B <- nrow(sim_nbinom)
threshold <- rep(1, d)
mean(apply(sim_nbinom, 1, function(X, y0 = threshold) {
  all( X >= y0)
}))
```

@@

```{r ch010-myM, include=FALSE, cache = TRUE}
m <- 10
```

@@

```{r ch050-quadlossECHO, echo=T, eval=F, cache=FALSE}
## Simulate random vectors equal to the sample size
N <- nrow(brca)
## create m random vectors and estimate correlation
simRho <- replicate(n = m, expr = {
  tmpSim <- rvec(n = N , R_X, nb_margins, cores = cores)
  bigsimr::cor_fast(x = tmpSim, method = type)
}, simplify = FALSE)
## find quadratic loss at each rep
quadLoss <- sapply(simRho, rags2ridges::loss, T = R_S, type = "quadratic")
```

```{r ch050-lossResults, echo=F, eval=T, cache=FALSE}
## quadLossRDS <- paste0( "results/brca", d, "_quadLoss.rds" )
frobeniusRDS <- paste0( "results/brca", d, "_frobenius.rds" )
if ( !file.exists( frobeniusRDS) )  {
    N <- nrow(brca)
    ## create m random vectors and estimate correlation
    simRho <- replicate(n = m, expr = {
        tmpSim <- rvec(n = N , R_X, nb_margins, cores = cores)
        bigsimr::cor_fast(x = tmpSim, method = type)
    }, simplify = FALSE)
    ## find quadratic loss at each rep
    ## simRho[[1]][1:5,1:5]
    ## R_S[1:5,1:5]
    ## rags2ridges::loss(simRho[[1]], R_S, type = 'frobenius')
    frobenius <- sapply(simRho, rags2ridges::loss, T = R_S, type = "frobenius")
    saveRDS(frobenius, frobeniusRDS)
}
frobenius <- readRDS( frobeniusRDS )
```

@@

```{r ch050-loss-summary, echo=TRUE, eval=T, cache=FALSE}
summary(frobenius)
```

>>>>

```{r ch050-LoadLib, include=FALSE, cache=FALSE}
devtools::load_all()
library(tidyverse)
library(bigsimr)

Sys.setenv(JULIA_NUM_THREADS = parallel::detectCores())
bs <- bigsimr_setup()
dist <- distributions_setup()

cores <- as.integer( parallel::detectCores() - 1 )
set.seed(10142020)
```

@@

```{r ch050-readBRCA, cache=FALSE}
d <- 1000
brca1000 <- example_brca %>%
  select(all_of(1:d)) %>%
  mutate(across(everything(), as.double)) # cor_fast currently doesn't work with integers
```

@@

With this goal in mind, we first estimate the desired correlation matrix using the fast implementation provided by `bigsimr`:

```{r ch050-estRhoBRCA, echo=TRUE, cache=FALSE}
# Estimate Spearman's correlation on the count data
type <- 'Spearman'
R_S <- bs$cor_fast(as.matrix(brca1000), bs[[type]])
```

Next, we estimate the marginal parameters. We use the method of moments (MoM) to estimate the marginal parameters for the multivariate negative binomial model. The marginal distributions are from the same probability family (NB) yet are heterogeneous in terms of the parameters probability and size $(p_i, n_i)$ for $i,\ldots,d$. The functions below support this estimation for later use in `rvec`.

<!-- 
TYPESTE MoM estimators here.
-->

```{r ch050-nbHelpers, echo=TRUE}
make_nbinom_margins <- function(sizes, probs) {
  margins <- lapply(1:length(sizes), function(i) {
    dist$NegativeBinomial(sizes[i], probs[i])
  })
  do.call(c, margins)
}

nbinom_mom <- function(x) {
  m <- mean(x)
  s <- sd(x)
  s2 <- s^2
  p <- m/s2
  r <- m^2 / (s2 - m)
  c(size = r, prob = p)
}
```

@@

```{r ch050-estMargins, echo = TRUE, eval = TRUE, cache=F}
nbinom_fit <- apply(brca1000, 2, nbinom_mom)
sizes <- nbinom_fit["size",]
probs <- nbinom_fit["prob",]

nb_margins <- make_nbinom_margins(sizes, probs)
```

@@

```{r ch050-cor-check, echo=TRUE}
# 1. Mapping step first
R_X <- bs$cor_convert(R_S, bs[[type]], bs$Pearson)

# 2a. Check admissibility
is_valid_correlation <- function(rho) {
  all(diag(rho) == 1) && matrixcalc::is.positive.semi.definite(rho)
}
is_valid_correlation(R_X)

# 2b. compute nearest correlation
if (!is_valid_correlation(R_X)) {
  R_Xtilde <- bs$cor_nearPD(R_X)
  ## Quantify the error
  targets <- R_X[lower.tri(R_X)]
  approximates <- R_Xtilde[lower.tri(R_Xtilde)]
  ## R_Xtilde[1:5, 1:5]
  ## norm( x = R_X - R_Xtilde, type = "F" )
  R_X  <- R_Xtilde
}
summary(targets - approximates)
```

While the exact $d \times d$ Spearman correlation matrix is not strictly admissible in our scheme (seen by the non-PSD result above), the approximation is of sufficient quality for most applications with a maximum pairwise of `r max( targets - approximates)` and average error of `r mean( targets - approximates)`.

```{r ch050-runBRCA-echo, echo=TRUE, eval=FALSE, cache=FALSE}
sim_nbinom <- bs$rvec(10000, R_X, nb_margins) 
colnames(sim_nbinom) <- colnames(brca1000)
```

```{r ch050-runBRCA, echo=FALSE, results=FALSE, cache=FALSE}
dir.create("results", showWarnings = FALSE, recursive = TRUE)
brcaSimRDS <- paste0("results/brca", d, "sim.rds")
if(!file.exists(brcaSimRDS)) {
  set.seed(44)
  n <- 10000
  sim_nbinom <- bs$rvec(n, R_X, nb_margins) 
  colnames(sim_nbinom) <- colnames(brca1000)
  saveRDS(sim_nbinom, brcaSimRDS)
}
sim_nbinom <- readRDS(brcaSimRDS)
```

@@

```{r ch050-makeSimDataFig, cache=FALSE}
if(!file.exists("fig/ch050-simDataFig.png")) {
  p <- GGally::ggpairs(
    data = as.data.frame(sim_nbinom[,example_genes[1:3]]),
    upper = list(continuous = GGally::wrap('cor', method = "spearman"))
  ) + theme_bw()
  ggsave("fig/ch050-simDataFig.png", plot = p, dpi = 300) 
}
```

```{r ch050-simDataFig, fig.cap='Simulated data for three selected high-expressing genes generally replicates the estimated data structure. The data do not exhibit outlying points, but do possess the desired Spearman correlations, central tendencies, and discrete values.'}
knitr::include_graphics('fig/ch050-simDataFig.png')
```

@@

```{r ch050-makeFigBRCA, cache=FALSE}
if(!file.exists("fig/ch050-figBRCA.png")) {
    ## estimate rho from sims and plot
    R_S_hat <- cor_fast(sim_nbinom, method = type)
    ## triplot
    ## png("fig/ch050-figBRCA.png", width = 4, res = 100, units = 'in')
    png("fig/ch050-figBRCA.png", res = 300, width = 5, height = 4, units = 'in')
    par(mfrow=c(1,3))
    plot(R_S[lower.tri(R_S)], R_S_hat[lower.tri(R_S_hat)],
         xlim = c(-1, 1), ylim = c(-1, 1),
         pch = 1, cex = 1, col = rgb(0, 0, 0, 0.05),
         xlab = "Target Correlation", ylab = "Estimated Correlation")
         ## main = "Spearman Cor. Matching")
    abline(a = 0, b = 1, col="red", lty="dashed")
    ## estimate prob,size from sims and plot
    nbinom_size_prob_hat <- apply(sim_nbinom, 2, nbinom_mom)
    plot(sizes, nbinom_size_prob_hat[1,],
         pch = 1, cex = 1, col = rgb(0, 0, 0, 1),
         xlab = "Target Size", ylab = "Estimated Size")
         ## main = "Marginal NB Size Matching")
    abline(a=0, b=1, col="red", lty="dashed")
    plot(log(probs), log(nbinom_size_prob_hat[2,]),
         pch = 1, cex = 1, col = rgb(0, 0, 0, 1),
         xlab = "Target Log Probability", ylab = "Estimated Log Probability")
         ## main = "Marginal NB Prob Matching")
    abline(a=0, b=1, col="red", lty="dashed")
    dev.off()
    par(mfrow=c(1,1))
}
```

```{r ch050-figBRCA, out.width='80%', fig.cap="Simulated random vectors from a multivariate negative binomial replicate the estimated structure from an RNA-seq data set. The dashed red lines indicated equality between estimated parameters from simulated data (vertical axes) and the specified target parameters (horizontal axes)."}
knitr::include_graphics('fig/ch050-figBRCA.png')
```

@@

```{r ch050-densityEvaluation, include=FALSE, cache=FALSE}
d <- ncol(sim_nbinom)
B <- nrow(sim_nbinom)
threshold <- rep(1, d)
pHat <- mean(apply(sim_nbinom, 1, function(X, y0 = threshold) {
  all(X >= y0)
}))
```

where ${\bf Y^{(b)} }$ is the $b^{th}$ simulated vector from a total of $B$ simulation replicates and $I(\cdot)$ is the indicator function. For example, we can estimate from our $B=10,000$ simulated vectors that the probability of all genes expressing (i.e., ${\bf y}_i \geq 1, \forall \; i$) is $`r pHat`$.

```{r ch050-densityEvaluationECHO, echo=TRUE, cache=FALSE}
d <- ncol(sim_nbinom)
B <- nrow(sim_nbinom)
threshold <- rep(1, d)
mean(apply(sim_nbinom, 1, function(X, y0 = threshold) {
  all( X >= y0)
}))
```

@@

```{r ch010-myM, include=FALSE}
m <- 10
```

@@

```{r ch050-quadlossECHO, echo=TRUE, eval=FALSE, cache=FALSE}
## Simulate random vectors equal to the sample size
N <- nrow(brca)
## create m random vectors and estimate correlation
simRho <- replicate(n = m, expr = {
  tmpSim <- bs$rvec(N , R_X, nb_margins)
  bs::cor_fast(tmpSim, bs[[type]])
}, simplify = FALSE)
## find quadratic loss at each rep
quadLoss <- sapply(simRho, rags2ridges::loss, T = R_S, type = "quadratic")
```

```{r ch050-lossResults, cache=FALSE}
frobeniusRDS <- paste0("results/brca", d, "_frobenius.rds")
if (!file.exists(frobeniusRDS)) {
    N <- nrow(brca1000)
    ## create m random vectors and estimate correlation
    simRho <- replicate(n = m, expr = {
        tmpSim <- bs$rvec(N , R_X, nb_margins)
        bs$cor_fast(tmpSim, bs[[type]])
    }, simplify = FALSE)
    frobenius <- sapply(simRho, rags2ridges::loss, T = R_S, type = "frobenius")
    saveRDS(frobenius, frobeniusRDS)
}
frobenius <- readRDS(frobeniusRDS)
```

@@

```{r ch050-loss-summary, echo=TRUE, eval=TRUE, cache=FALSE}
summary(frobenius)
```

### 2021-02-24 index.Rmd

```{r indexSetup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(eval = TRUE, 
                      echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      cache = TRUE)

library(tidyverse)
library(knitr)
library(printr)
library(rmarkdown)
library(bookdown)

@@

```{r serveHelper, include=FALSE, eval=FALSE}
## rmote::start_rmote()
## library(ggplot2); qplot(mpg, wt, data=mtcars, colour=cyl)
## bookdown::render_book("index.Rmd", "bookdown::pdf_book", output_dir = "pdf")
## bookdown::preview_chapter("index.Rmd")
rmarkdown::render_site(encoding = 'UTF-8')
bookdown::serve_book()
```

```{r countWords, include=FALSE, eval=FALSE}
rmdFiles <- dir(pattern = 'Rmd$', ignore.case= T)
## rmdFiles <- rmdFiles[-grep('^_', rmdFiles)]
rmdFiles <- rmdFiles[-grep('references', rmdFiles)]
## rmdFiles <- rmdFiles[-grep('README', rmdFiles)]
## rmdFiles <- rmdFiles[-grep('article', rmdFiles)] ## after generating a single document
## wordcountaddin::word_count( rmdFiles[11] )
(counts <- sapply( rmdFiles, wordcountaddin::word_count, simplify = T ))
sum(counts)
```

>>>>

```{r indexSetup, include=FALSE, cache=FALSE}
library(tidyverse)
library(knitr)
library(printr)
library(rmarkdown)
library(bookdown)

@@

```{r serveHelper, include=FALSE, eval=FALSE}
# rmote::start_rmote()
# library(ggplot2); qplot(mpg, wt, data=mtcars, colour=cyl)
# bookdown::render_book("index.Rmd", "bookdown::pdf_book", output_dir = "pdf")
# bookdown::preview_chapter("index.Rmd")
rmarkdown::render_site(encoding = 'UTF-8')
bookdown::serve_book()
```

```{r countWords, include=FALSE, eval=FALSE}
rmdFiles <- dir(pattern = 'Rmd$', ignore.case= T)
# rmdFiles <- rmdFiles[-grep('^_', rmdFiles)]
rmdFiles <- rmdFiles[-grep('references', rmdFiles)]
# rmdFiles <- rmdFiles[-grep('README', rmdFiles)]
# rmdFiles <- rmdFiles[-grep('article', rmdFiles)] ## after generating a single document
# wordcountaddin::word_count( rmdFiles[11] )
(counts <- sapply( rmdFiles, wordcountaddin::word_count, simplify = T ))
sum(counts)
```
